Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-10-02 paper.tex
edits up to results section

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  The other component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps create a need for larger floating
inventory buffers in order to avoid material shortages.  Another aspect of
this effect is that a single material object can only move between a single
pair of facilities per time step. Larger time steps result in longer minimum
bounds on time to traverse paths between facilities. Although direct measures
of this effect are difficult, some of its consequences (e.g. higher standing
inventories during shortages) can be quantified by comparing scenarios with
different time step durations (e.g. cases MF and QF).

There are also performance differences between the two modeling paradigms.  A
single 200 year simulation with several hundred fleet-based reactors takes
about 2 seconds to run on a typical desktop machine where the same simulation
using individual reactor facilities takes about 20 seconds.  This order of
magnitude difference is an important tradeoff to keep in mind when doing fuel
cycle analysis in general, particularly when considering optimization or
sensitivity studies where running hundreds of thousands of simulations may be
desirable.  The time step duration has less impact on performance because the
amount of work being done in the simulation is the same; in the monthly time
step cases most facilities (i.e. all reactors) are only perform actions every
refueling cycle rather than every time step.

>>>>

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  However, another component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps create a need for larger floating
inventory buffers in order to avoid material shortages.  Another aspect of
this effect is that at least one time step is required for a material object to move between two facilities. Larger time steps result in longer minimum
bounds on time to traverse paths between facilities. Although direct measures
of this effect are difficult, some of its consequences (e.g. higher standing
inventories during shortages) can be quantified by comparing scenarios with
different time step durations (e.g. cases MF and QF).

There are also performance differences between the two modeling paradigms.  A
single 200 year simulation with several hundred fleet-based reactors takes
about 2 seconds to run on a typical desktop machine where the same simulation
using individual reactor facilities takes about 20 seconds.  This order of
magnitude difference is an important tradeoff to keep in mind when doing fuel
cycle analysis in general, particularly when considering optimization or
sensitivity studies where running thousands or millions of simulations may be
desirable.  The time step duration has less impact on performance because the
amount of work being done in the simulation is similar; in the monthly time
step cases most facilities (i.e. all reactors) only perform actions every
refueling cycle rather than every time step.
